Comparing Generative Pre-trained Transformer Models to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-04-13 manuscript/manuscript.tex
Done editing adult perf

Across the adult age ranges, GPT-4 had the highest PCCC values at or above 0.59, followed by GPT-3.5, while InterVA-5 and InSilicoVA had the lowest PCCC values at or below 0.46. GPT-4 PCCC was between 0.61 and 0.72 for young to middle age ranges between 10 to 49 years, and at or below 0.62 for older age ranges between 50 to 69 years (Figure \ref{fig-perf-agerange-adult}). Each age range had between 135 (4\%) to 575 (15\%) records with the 65-69 year age range having the most records (n=575, 15\%), while other age ranges ranged from 4-9\% of all records. Figure \ref{fig-perf-cod-adult} shows the highest and lowest performing models by PCCC across 18 adult CODs. The highest PCCC values were GPT-4 between 0.36 to 0.99 for 12 of 18 CODs (GPT-3.5 within 0.05 for 7 of the 15 CODs), InSilicoVA for road and transport injuries at 0.84 (InterVA-5 close at 0.81), GPT-3.5 for ill-defined at 0.57,and InterVA-5 for ischemic heart disease at 0.61 (GPT-4 close at 0.57). The lowest PCCC values were InterVA-5 between 0 to 0.79 for 11 of 18 CODs (InSilicoVA within 0.05 for 3 of the 11 CODs), InSilicoVA between 0.01 to 0.41 for 6 of 18 CODs (InterVA-5 within 0.05 for other cardiovascular diseases and diarrhoeal diseases), and GPT-3.5 for ischemic heart disease at 0.49 (InSilicoVA within 0.05) and road and transport injuries at 0.59. Suicide and chronic respiratory diseases had the largest differences between minimum and maximum PCCC at 1 and 0.81 respectively. Maternal conditions, road and transport injuries, and tuberculosis had higher PCCC values above or at 0.67 (with the exception of GPT-3.5 for road and transportation injuries as an outlier), where there were small differences between minimum and maximum PCCC ranging from 0.12 to 0.19. Ischemic heart disease and unspecified infections had lower PCCC values below or at 0.6, where there were small differences between minimum and maximum PCCC ranging at 0.02 and 0.17 respectively. Each adult COD had between 3 (\textless 1\%) to 799 records (21\%) with malaria (n=799, 21\%), other noncommunicable diseases (n=414, 11\%), and unspecified infections (n=389, 10\%) making 42\% of all records, suicide having less than 1\% of records, and other CODs between 1-9\% of all records. GPT-4 had the highest PCCC for both male and female records at 0.72 and 0.7 respectively, while all other models had PCCC at or below 0.63 (see Figure \ref{fig-perf-sex-adult} in Appendix \ref{app-perf-misc}).

@@

\section{Case Study Methods}

>>>>

Across the adult age ranges, GPT-4 had the highest PCCC values at or above 0.59, followed by GPT-3.5, while InterVA-5 and InSilicoVA had the lowest PCCC values at or below 0.46. GPT-4 PCCC was between 0.61 and 0.72 for young to middle age ranges between 10 to 49 years, and at or below 0.62 for older age ranges between 50 to 69 years (Figure \ref{fig-perf-agerange-adult}). Each age range had between 135 (4\%) to 575 (15\%) records with the 65-69 year age range having the most records (n=575, 15\%), while other age ranges ranged from 4-9\% of all records. Figure \ref{fig-perf-cod-adult} shows the highest and lowest performing models by PCCC across 18 adult CODs. The highest PCCC values were GPT-4 between 0.36 to 0.99 for 10 of 17 CODs (GPT-3.5, InterVA-5, and InSilicoVA within 0.05 for 3, 1, and 1 of these 10 CODs respectively), InSilicoVA for tuberculosis and road and transport injuries at 0.72 and 0.84 (GPT-4 and InterVA-5 within 0.05), and GPT-3.5 between 0.44 and 0.94 for 5 of 17 CODs (GPT-4 within 0.05 for malaria). The lowest PCCC values were InterVA-5 between 0 to 0.79 for 8 of 17 CODs (InSilicoVA and GPT-3.5 within 0.05 for cancers and unspecified infections respectively), InSilicoVA between 0.01 to 0.41 for 6 of 17 CODs (InterVA-5 within 0.05 for other cardiovascular diseases and diarrhoeal diseases), and GPT-3.5 at 0.39 for tuberculosis and 0.54 for road and transport injuries. Chronic respiratory diseases had the largest differences between minimum and maximum PCCC at 0.93. Maternal conditions, and road and transport injuries, had higher PCCC values above or at 0.67 (with the exception of GPT-3.5 for road and transportation injuries as an outlier), where there were small differences between minimum and maximum PCCC below or at 0.2. Ischemic heart disease and unspecified infections had PCCC values below or at 0.65, where there were small differences between minimum and maximum PCCC at or below 0.14. Each adult COD had between 27 (\textless 1\%) to 799 records (21\%) with malaria (n=799, 21\%), other noncommunicable diseases (n=414, 11\%), and unspecified infections (n=389, 10\%) making 42\% of all records, and other CODs between 1-9\% of all records. GPT-4 had the highest PCCC for both male and female records at 0.65 and 0.64 respectively, while all other models had PCCC at or below 0.6 (see Figure \ref{fig-perf-sex-adult} in Appendix \ref{app-perf-misc}).

@@

\section{Case Study Methods}\label{app-methods}
